\begin{coverpage}{Zulip - a Communication Platform for Collaborations}
{
\subsubsection*{What is Zulip?}
\textbf{Zulip} stands out as a dynamic and free communication platform, offering a structured and organized way of messaging, similar to Slack.

[Zulip screenshot]

Currently, our Zulip instance is self-hosted by the Computational Omics Laboratory at Pusan National University. This means it is completely \textbf{free}, and has no bulls**t, e.g., a 90-day message limit. Moreover, it comes with a Korean-to-English translator (based on Papago translator), which automatically translates your Korean messages into English, so that you can even use Korean to communicate with English-speaking collaborators.

[Some comparison chart]

Thus we strongly recommend you to use Zulip as a robust and free alternative to other means of communication between institutions for your research. We hope that Zulip becomes your favorite go-to platform for effective, efficient collaboration within our community and beyond.

\subsection*{1. How to Register and Log In}
First, you need to install a mobile or desktop app. Zulip is accessible on various platforms, including Windows, macOS, Linux, Android, and iOS. Visit \textbf{\texttt{https://zulip.com/apps/}} to install the suitable one. After that, open the installed app and follow the below instructions:
\begin{itemize}[leftmargin=0.5cm, rightmargin=0.5cm]
\item \textbf{Enter the Zulip Server URL:} Open Zulip app, and type \\ \textbf{\texttt{https://zulip.scsok.io/}}.
\item \textbf{Sign Up:} Click \textbf{\texttt{Sign Up}} and use your email or GitHub account to register.
\item \textbf{Log In:} Enter your credentials to sign in.
\end{itemize}
\subsection*{2. Sending Messages in Streams}
\begin{itemize}[leftmargin=0.5cm, rightmargin=0.5cm]
\item \textbf{Join a Stream:} Find and join relevant streams.
\item \textbf{Send Messages:} Use the input box to type messages.
\item \textbf{Change Notifications Settings:} Customize notification settings in 'Settings' > 'Notifications'.
\end{itemize}
\subsection*{3. Sending Direct Messages to Each Other}
\begin{itemize}[leftmargin=0.5cm, rightmargin=0.5cm]
\item \textbf{Find a User:} Click 'New Private Message'.
\item \textbf{Enter Recipient:} Add the recipient's name.
\item \textbf{Compose and Send:} Write and send your message.
\end{itemize}
\subsection*{5. How to Create a New Stream for Your Collaboration}
\begin{itemize}[leftmargin=0.5cm, rightmargin=0.5cm]
\item \textbf{Create Stream:} Access 'Streams' > 'Create Stream'.
\item \textbf{Set Details:} Name your stream and add a description.
\item \textbf{Add Members:} Invite collaborators.
\item \textbf{Configure:} Adjust privacy and settings.
\end{itemize}
}
\end{coverpage}
